Online resolving assembly graph by ONT long reads data

### 2019-07-22 paper.tex
method modify

\subsection*{Multiplicity estimation}
Due to the possible divergence of sequencing coverage relative to the real abundance of sequences, especially the short ones, an optimization step is carried out to alleviate this issue as shown in Supplementary Note \ref{supp:note2}. The coverage measures of nodes (which represent contigs) are spread throughout the graph via connecting edges for comparisons and calibrations prior to the multiplicity estimation. 
%The assignment of coverage to edges is also helpful to induce multiplicity of nodes in later step.

@@

%\subsection*{Untangling assembly graph by stream of nanopore data.}
\subsection*{Building bridges in real-time}
Bridge is the data structure designed for tracking the possible connections between two unique contigs (anchor nodes in the assembly graph).
Here we take advantage of the graph topology and nodes' multiplicity information to employ a dynamic bridging mechanism.
This procedure considers the dynamic changing of multiplicity property for each node, meaning that a $n$-times repetitive node can become a unique node at certain time point when its $(n-1)$ occurrences have been already identified and assigned in appropriate unique paths.

Other than \npscarf{}, a bridge in \npgraph{} has several completion levels. A bridge is only created with at least one unique contig as an end, if so it has level 1 of completion. If both ends are identified, the level is 2. The number is greater than that only if paths connecting two ends are found. A bridge is known as fully complete (level 4) if there is only one unique linking path left. Given a bridge with 2 ends, a path finding algorithm (described in next section) is invoked to find all candidate paths. Each of these paths is given a score of alignment-based likelihood which are updated immediately as long as there is an appropriate long read being generated by the sequencer. As more nanopore data arrives, the divergence between candidates' score becomes greater and only the best ones are voted for the next round.

Whenever a bridge becomes complete thanks to the voting system, the assembly graph is \emph{transformed} or \emph{reduced} by replacing its unique path by an composite edge and removing any unique edges (edges coming from unique nodes) along the path. The assembly graph would have at least one edge less than the original after the reduction. The nodes located on the reduced path, other than 2 ends, also have their multiplicities subtracted by one and the bridge is marked as finally resolved without any further modifications.

\subsection*{Path finding algorithm}

We implement a modified stack-based version utilizing Dijkstra's shortest path finding algorithm~\cite{Dijkstra1959} to reduce the search space when using Depth-First Search (Supplementary Note~\ref{supp:note3}).

\subsection*{Result extraction and output}
\npgraph{} reports assembly result in real-time by decomposing the assembly graph into a set of longest straight paths (LSP), each of the LSP will present a contig for the result (Supplementary Note \ref{supp:note4}).

>>>>

\subsection*{Multiplicity estimation}
Due to the possible divergence of sequencing coverage relative to the real abundance of sequences, especially for the shorter contigs, an optimization step is carried out to alleviate this issue as shown in Supplementary Note \ref{supp:note2}. The coverage measures of nodes (which represent contigs) are propated throughout the graph via connecting edges for calibrations prior to the multiplicity estimation. 
%The assignment of coverage to edges is also helpful to induce multiplicity of nodes in later step.

@@

%\subsection*{Untangling assembly graph by stream of nanopore data.}
\subsection*{Building bridges in real-time}
Bridge is the data structure designed for tracking the possible connections between two anchored nodes (of unique contigs) in the assembly graph.
The real-time bridging method considers the dynamic aspect of multiplicity measures for each node, meaning that a $n$-times repetitive node might become a unique node at certain time point when its $(n-1)$ occurrences have been already identified in other distinct unique paths. 
Furthermore, the streaming fashion of this method allows the bridge constructions to be carried out progressively so that assembly decisions are made immediately after having sufficient supporting data.
The detail implementation is addressed in Supplementary Note~\ref{supp:note3}.

A bridge in \npgraph{} has several completion levels. When created, it must be rooted from an \emph{anchor node} which represents a unique contig (level 1). A bridge is known as fully complete (level 4) if and only if there is a unique path connecting its two anchor nodes from two ends.

\subsection*{Path finding algorithm}
Given a bridge with 2 anchors, a path finding algorithm is invoked to find all candidate paths between them. Each of these paths is given a score of alignment-based likelihood which are updated immediately as long as there is an appropriate long read being generated by the sequencer. As more nanopore data arrives, the divergence between candidates' score becomes greater and only the top-scored ones are kept for the next round.
We implement a modified stack-based version utilizing Dijkstra's shortest path finding algorithm~\cite{Dijkstra1959} to reduce the search space when using Depth-First Search (Supplementary Note~\ref{supp:note4}).

It is worth noting that due to dead-ends, there not always exist a path in the assembly graph connecting two anchors as suggested by the alignments. In this case, if enough long reads coverage (20X) are met, a consensus module is invoked and the resulting sequence is used for a \emph{pseudo} edge being created.
\subsection{Resolve by reducing the graph complexity}

Whenever a bridge is finished (with only one path left), the assembly graph is \emph{transformed} or \emph{reduced} by replacing its unique path by an composite edge and removing any unique edges (edges coming from unique nodes) along the path. The assembly graph would have at least one edge less than the original after the reduction. The nodes located on the reduced path, other than 2 ends, also have their multiplicities subtracted by one and the bridge is marked as finally resolved without any further modifications.

\subsection*{Result extraction and output}
\npgraph{} reports assembly result in real-time by decomposing the assembly graph into a set of longest straight paths (LSP), each of the LSP will present a contig for the result (Supplementary Note \ref{supp:note5}).
